Kindly Bent to Free Us

### 2019-06-28 kindly/motivation.tex
Explain the pool example a bit further.

\subsection{Local data-structures with regions}

We have seen that borrows are limited to a region. In the various examples so
far, such regions didn't appear in the type signature explicitly.
We know present an example implement of a \texttt{Pool} module
that uses such region to enforce linearity of the objects in the pool.

@@

let create spawn =
  { spawn ; queue = Queue.create () }
let consume f c = Queue.iter f c.queue

let use { spawn ; queue } f =
  let o = match Queue.pop &queue with
    | Some x -> x
    | None () -> spawn ()
  in
  let r = f &!o in
  Queue.push o &queue;
  r
\end{lstlisting}
    \caption{Implementation}
  \end{subfigure}~
  \begin{subfigure}[t]{.55\linewidth}
\begin{lstlisting}
type ('a: 'k) pool : 'k

create : (unit -> 'a) -> 'a pool
consume : ('a -> unit) -> 'a pool -> unit
use : ('a_1:'k_1),('a_2:'k_2),('k_1 < lin_r),('k_2 < lin_r) =>
  &('a_1 pool) -> (&!(aff_r,'a_1) -{'k}> 'a_2) -{'k_1}> 'a_2
\end{lstlisting}
    \caption{Interface}

@@

(* Use the pool in queries. *)
let create_user name =
  Pool.use &pool (fun connection ->
    Db.insert "users" [("name", name)] connection)
\end{lstlisting}
    \caption{Example of use}
  \end{subfigure}

  \caption{The \texttt{Pool} module}
\end{figure}

>>>>

\subsection{Local data-structures with regions}
\label{tuto:pool}

We have seen that borrows are limited to a region. In the various examples so
far, such regions didn't appear in the type signatures explicitly.
We know present an pool implementation which can store linear objects
and such region to enforce proper uses of the objects.
The implementation and interfaces of the \texttt{Pool} module
are shown in \cref{impl:pool,intf:pool}. A simple example of use
is shown in \cref{ex:pool}.

A pool is parameterized by its content. Additional, the kind of the pool
depends on the content: Linear content implies
a linear pool, while unrestricted content implies an unrestricted pool.
The implementation uses a concurrent queue
as a bag of elements with atomic add and remove operations.
The \lstinline/Pool.create/ and \lstinline/Pool.consume/ allow to build or destroy
the pool by providing the necessary components for the elements
of the loop.
The main function of interest is the \lstinline/Pool.use/ function, which
allow to use an mutable borrow of the object. The function proceeds by removing
the element from the pool (or creating a fresh element),
passing it down to the \lstinline/f/ function, and then adding
it back to the pool.
We explicitly delimit the region on \cref{line:pool:region} to ensure that
the \lstinline/f/ function will not be able to capture \lstinline/&!o/.

@@

let create spawn =
  { spawn ; queue = CQueue.create () }
let consume f c = CQueue.iter f c.queue

let use { spawn ; queue } f =
  let o = match CQueue.pop &queue with
    | Some x -> x
    | None () -> spawn ()
  in
  let r = {| f &!o |} in(*@\label{line:pool:region}*)
  Queue.push o &queue;
  r
\end{lstlisting}
    \caption{Implementation}
    \label{impl:pool}
  \end{subfigure}~
  \begin{subfigure}[t]{.55\linewidth}
\begin{lstlisting}
type ('a: 'k) pool : 'k

create : (unit -> 'a) -> 'a pool
consume : ('a -> unit) -> 'a pool -> unit
use : ('a_1:'k_1),('a_2:'k_2),('k_1 < lin_r),('k_2 < lin_r) =>
  &('a_1 pool) -> (&!(aff_r,'a_1) -{'k}> 'a_2) -{'k_1}> 'a_2
\end{lstlisting}
    \caption{Interface}
    \label{intf:pool}

@@

(* Use the pool in queries. *)
let create_user name =
  Pool.use &pool (fun connection ->
    Db.insert "users" [("name", name)] connection)
\end{lstlisting}
    \caption{Example of use}
    \label{ex:pool}
  \end{subfigure}

  \caption{The \texttt{Pool} module}
  \label{fig:pool}
\end{figure}
